### 2015-05-14 corral.tex
final updates in corral section

\noindent\textbf{Race-Checking Instrumentation }
%
To detect data races with \corral, \whoop outputs a Boogie program instrumented with a simple, but effective encoding of race checks. Whenever there is a write access to a shared variable $s$, \whoop instruments the program as follows:
%
\begin{boogie}
s = e;         // original write
yield;         // allow for a context-switch
assert s == e; // check written value
\end{boogie}
%
Likewise, whenever there is a read access to $s$, \whoop instruments the program as follows:
%
\begin{boogie}
x = s;         // original read
yield;         // allow for a context-switch
assert x == s; // check read value
\end{boogie}

@@

\noindent\textbf{Sound Partial-Order Reduction }
%
By default, and assuming no race-freedom guarantees, \whoop instruments a \texttt{yield} statement (which denotes a potential nondeterministic context-switch) after each shared memory access of each entry point, and after every lock and unlock operation.\footnote{We acknowledge that in the presence of data races and relaxed memory, even considering all interleavings of shared memory accesses may be insufficient to find all bugs.} \whoop then sends this instrumented program to \corral, which leverages sequentialization to explore all possible thread interleavings up to a pre-defined bound. Our approach to accelerating \corral is simple and yet effective: if, thanks to \whoop's analysis, we know that a given statement that accesses shared memory cannot be involved in a data race, we do not instrument a \texttt{yield} after this statement.  This is a form of \emph{partial order reduction}~\cite{DBLP:books/sp/Godefroid96}, and reduces the number of context switches that \corral must consider in a \emph{sound} manner: there is no impact on the bugs that will be detected.  This is because each access for which a \texttt{yield} is suppressed is guaranteed to be protected by some lock (a consequence of lockset analysis).  If the access is a write, its effects are not visible by the other entry point until the lock is released.  If the access is a read, the value of the shared location cannot change until the lock is released.  The fact that a \texttt{yield} is placed after each unlock operation suffices to take account of communication between entry points via the shared memory location.

\ADComment{Don't say ``we regard races as bugs''.  After POR part, emphasize that this idea can be used to accelerate Corral for arbitrary bug-finding in concurrent programs.  Then make it clear that in the remainder of the paper we regard races as bugs and use corral solely to look for races.}

\noindent\textbf{Granularity of Context-Switches }
%
We have implemented two different levels of context-switch instrumentation granularity:

@@

In this work, we use \corral to analyze individual pairs of entry points.
We do not employ any abstraction to model additional threads, because we want \corral to report only true races. Because we only analyze pairs, though, \corral will miss races that require more than two threads to manifest.
We could extend our setup so that more than two threads are considered by \corral, though because the number of threads that may be launched by an OS kernel is unknown in general, we are inevitably limited by some fixed maximum thread count.

>>>>

\noindent\textbf{Race-Checking Instrumentation }
%
To detect data races with \corral, \whoop outputs a Boogie program instrumented with a simple, but effective encoding of race checks. Whenever there is a write access to a shared variable $s$, \whoop instruments the program as follows:
%
\begin{boogie}
s = e;         // original write
yield;         // allow for a context-switch
assert s == e; // check written value
\end{boogie}
%
Likewise, whenever there is a read access to $s$, \whoop instruments the program as follows:
%
\begin{boogie}
x = s;         // original read
yield;         // allow for a context-switch
assert x == s; // check read value
\end{boogie}
%
A \texttt{yield} statement denotes a nondeterministic context-switch, and is used by \corral to guide the sequentialization.

@@

In this work, we use \corral to analyze individual pairs of entry points. We do not use any abstraction to model additional threads, as we want \corral to report only true races. Because we only analyze pairs, though, \corral will miss races that require more than two threads to manifest. We could extend our setup so that more than two threads are considered by \corral, but because the number of threads that an OS kernel might launch is unknown in general, we are inevitably limited by some fixed maximum thread count.

\noindent\textbf{Sound Partial-Order Reduction }
%
By default, and assuming no race-freedom guarantees, \whoop instruments a \texttt{yield} after each shared memory access of each entry point, and after every lock and unlock operation.\footnote{We acknowledge that in the presence of data races and relaxed memory, even considering all interleavings of shared memory accesses may be insufficient to find all bugs.} \whoop then sends this instrumented program to \corral, which leverages sequentialization to explore all possible thread interleavings up to a pre-defined bound. Our approach to accelerating \corral is simple and yet effective: if, thanks to \whoop's analysis, we know that a given statement that accesses shared memory cannot be involved in a data race, we do not instrument a \texttt{yield} after this statement.  This is a form of \emph{partial order reduction}~\cite{DBLP:books/sp/Godefroid96}, and reduces the number of context switches that \corral must consider in a \emph{sound} manner: there is no impact on the bugs that will be detected.  This is because each access, for which a \texttt{yield} is suppressed, is guaranteed to be protected by some lock (a consequence of lockset analysis).  If the access is a write, its effects are not visible by the other entry point in the pair until the lock is released.  If the access is a read, the value of the shared location cannot change until the lock is released.  The fact that a \texttt{yield} is placed after each unlock operation suffices to take account of communication between entry points via the shared memory location.

We have implemented two different levels of context-switch instrumentation granularity:

@@

Our partial-order reduction can be used to accelerate \corral for arbitrary bug-finding in concurrent programs. In the remainder of this paper, though, we regard races as bugs and use \corral solely to look for races.
